### 2018-10-31 tdd-in-scrum.tex
talk: renamed the tex into the name of the talk

\documentclass{f4_beamer_metropolis}

\title{Test-Driven Development in Scrumentwicklung}
\subtitle{PMQM WS18/19}
\author{Dennis Grabowski, B.Sc}
\date{17.11.2018}

\begin{document}

% Cannot be moved into the class
\begin{frame}{Inhaltsverzeichnis}
    \tableofcontents[hideallsubsections]
  \end{frame}

\begin{frame}[standout]
  Offene Fragen?
\end{frame}

\end{document}